Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function

### 2023-01-04 manuscript/manuscript.Rmd
revice manuscript

Microbiomes mediate a variety of important functions in ecosystems, and there
is great interest in understanding how attributes of microbiomes may influence
variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019].
Biodiversity-ecosystem function relationships have been documented for a
variety of macroorganismal communities. For example, plant species richness
correlates with productivity and marine community diversity correlates with
production, consumption, and nutrient release 
[@cardinale2006; @hooper2012; @gamfeldt2015]. However, we have been unsuccessful
in documenting similar relationships for microbial communities. While there is
some evidence that microbiomes from different ecosystems display different
functional rates [@balser2005; @strickland2009], adding microbiome attributes to
ecosystem models adds very little to total variance explained [@graham2016].
Therefore, we still lack a detailed understanding of what
functions are altered by microbiome composition and what attributes of the
microbiome explain this relationship.

There are several reasons for this difficulty. One reason is that it is
challenging to manipulate microbiome composition and diversity directly.
Experimental manipulations of community composition and diversity have been
central to documenting the relationships between ecosystem functions and the
biodiversity of plants and animals [@hooper2012], but this is seldom possible
with microbiomes. Most microorganisms cannot yet be cultured and most available
methods for manipulation of microbiomes (e.g., antibiotics, dilution, etc.) are
not sufficiently specific to manipulate individual microbial lineages. In
addition, many studies have looked for relationships between an ecosystem
function and the abundance or diversity of a microbial gene that codes for an
enzyme or enzyme-subunit known to be involved in that function. However, it is
difficult to identify the “right” marker gene a priori and for most functions
in most ecosystems there is no correlation between the abundance of a marker
gene and the rate of the corresponding process [@rocca2015]. Lastly,
microbiome-ecosystem function relationships have been difficult to document
because microbiomes often covary with other drivers of variation in ecosystem
functions, such as environmental conditions or resources. Such covariation can
make it impossible to determine whether microbiomes directly contribute to
functional variation. Understanding the degree to which microbiomes directly
influence functional variation is crucial if scientists are to develop
microbiome-focused methods for altering or maintaining ecosystem functions.

One way to overcome these challenges is to manipulate the relationship between
microbiome biodiversity and ecosystem function using artificial ecosystem selection 
[@swenson2000; @blouin2015; @panke-buisse2015; @sanchez2021]. 
Microbiomes have been shown to respond to selection at the ecosystem-trait
level [@swenson2000], and some ecosystem-scale microbiome traits have been
shown to be transferrable through microbiome inoculations [@panke-buisse2015].
Therefore, applying artificial selection to whole microbiomes could be a novel
approach for describing a more detailed microbiome-ecosystem function
relationship. This approach works by subsampling a microbiome to inoculate
multiple replicate ecosystems that vary in microbiome composition and then
measuring the rate of ecosystem function for each replicate. The investigator
then selects the replicates with the highest (or lowest) rates and subsamples
those ecosystems to passage them to a new set of ecosystems. By holding
environmental conditions relatively constant among the replicates and from
passage to passage (or by using isogenic hosts, in the case of host-associated
microbiomes), investigators can determine whether variation in the microbiome
contributes to variation in the rate of ecosystem function independent of the
environment.

A valuable feature of this approach is that it can be used to estimate how much
of the variation in ecosystem function is due to microbiome variation.  By
calculating the slope of the regression between the "donor" ecosystems selected
in one passage and the "recipient" ecosystems inoculated by those donors in the
next passage, we can quantify an estimate of "ecosystem heritability"
[@goodnight2000; @blouin2015]. The concept of heritability has been borrowed
from quantitative genetics where heritability is defined as *the ratio of
additive genetic variance to the total phenotypic variance* and is commonly
estimated as *the slope of the linear regression between offspring phenotype
and parental phenotype* [@bell1977; @falconer1996]. Here we apply this concept
to donor and recipient communities in an artificial ecosystem selection
experiment to estimate *the variance in ecosystem function due to microbiome
variance*. We do not assume that environmental variation is zero, instead we
quantify the proportion of variation due to microbiome effects. If the slope of
the regression is zero, we would conclude that environmental variation among
the replicates and technical error are the primary drivers of variation in
ecosystem function. On the other hand, a significant slope between donors and
recipients would demonstrate that microbiome variation generates variation in
the rate of an ecosystem function. We then use our artificially selected
microbiomes to investigate what aspects of microbiome variation explain this
response. By "microbiome variation" we mean aspects of microbial diversity
commonly measured by microbial ecologists, such as the relative abundance of
taxonomic groups inferred from amplicon sequence variants. With this approach,
we can not only determine whether an ecosystem function is influenced by
microbiome variation, but also estimate how much of the variation and which
taxa are associated with higher rates of ecosystem function.

We applied artificial ecosystem selection to soil microbial communities by
selecting on soil methane oxidation rate. We chose this function because
methane is a globally important greenhouse gas and methane oxidation by soil
bacteria is the primary biological sink for atmospheric methane
[@kirschke2013]. In addition, there is evidence that soil methane oxidation
rate may vary with microbiome composition. For example, soil microbiome
phylogenetic variation is a strong predictor of methane oxidation in forests
and pastures of the Brazilian Amazon as well as soils of the Congo Basin
[@meyer2020; @meyer2020a]. Finally, methanotrophy is one of the most deeply
conserved microbial physiologies and is represented in a narrow range of taxa
and so the taxonomic composition of the microbiome is more likely to be
associated with the rate of methane oxidation than other broader or more
shallowly conserved functions [@schimel1998; @martiny2013]. This suggests that
microbiome variation might drive variation in methane oxidation in soils. One
other study has estimated ecosystem heritability for an environmental
microbiome. Blouin et al. [@blouin2015] selected for water microbiomes with
low carbon dioxide respiration rates and estimated a heritability of 52%.
However, this appeared to be driven primarily by biomass, not microbiome
composition. Here we want to build on this work by selecting for higher rates
of a more narrow function that is less likely to be driven by total biomass.
That way we can begin to assess whether variation in microbiome composition is
important for understanding variation in ecosystem functions.

@@

We estimated the ecosystem heritability of methane oxidation rate, which is the
proportion of variance in CH~4~ oxidation rate due to variation in the
microbiome, as the regression of mid-recipient on mid-donor [@falconer1996].
The mid-donor was the mean for all three selected jars and the mid-recipient was
the mean methane oxidation rate for the twelve jars produced by those three
donors. First, we tested if there was a difference of slopes between the
positive and neutral treatments. To test this, we fit two nested models with and
without the treatment term and compared them using an F-test with the 'anova'
function. We then present the slopes for each treatment, which represented the
ecosystem heritability or the proprotion of total variance in methane oxidation
rate due to microbiome variance.

First, we tested if there was an effect of treatment by comparing
nested models with and without treatment using an F-test with the 'anova'
function. We then fit the full model to estimate the slope of the relationship
between recipient and donor for the Positive and Neutral treatments.

@@

@@

While it is well understood that the microbiome mediates a variety of ecosystem
functions, a fundamental question in microbial ecology is whether we need to
understand variation in the composition of microbiomes to understand variation
in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this
question, we performed artificial ecosystem selection on soil microbiomes by
selecting for ecosystems with high methane oxidation rates. We found that there
was a significant increase in methane oxidation rate in the Positive selection
treatment relative to the Neutral selection treatment. This demonstrates that
passaging the microbiome through artificial ecosystem selection resulted in a
shift in the mean methane oxidation rate of the soil microbiome. 
This suggests that variation in the microbiome can generate variation in
ecosystem functions independent of the environment.

We quantified how much of the variation in methane oxidation rate was due to
variation in the microbiome by regressing the mean methane oxidation rate of
the 12 recipient jars against the mean methane oxidation rate of the three
donor jars selected to inoculate the recipients. We observed a heritability of
`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,1)`\% 
for the Positive selection treatment. This indicates that 
`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,1)`\% 
of the variation in methane oxidation rate can be attributed to variation in the
microbiome. This is comparable to the ecosystem heritability of carbon dioxide
emissions from water [@blouin2015]. Notably, we observed a negative heritability
in the Neutral treatment (`r myround(coefficients(herit_model)[2]*100, 1)`\%).
Negative heritability estimates are not uncommon in quantitative genetics
[@steinsaltz2020]. This can occur for random selection controls in artificial
selection experiments and may indicate that methane oxidation rate is negatively
correlated with other traits associated with persistence in jars. The
heritability we observed in the Positive treatment indicates that a large
portion of the variation in methane oxidation rate is due to variation in the
microbiome in our experimental system.

@@

>>>>

Microbiomes mediate a variety of important ecosystem functions. This suggests
that understanding variation in microbiome composition might be important for
predicting the rates of ecosystem functions 
[@conrad1996; @schimel1998; @crowther2019]. 
Many studies have attempted to document such a relationship
[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009; @wagg2014; @wagg2019].
However, it is difficult to isolate the effect of variation in microbiome
composition from other drivers of variation in ecosystem function. For example,
microbiome manipulations often confound variation in diversity, composition, and
abundance. Likewise, comparative studies must control for abiotic variation
statistically to identify the direct effect of microbiome variation. This
approach relies on measuring the relevant abiotic variables and cannot account
for unmeasured environmental variation. Here, we overcome these limitations by
using a selection approach to estimate the degree to which an ecosystem function
varies with microbiome composition.

There have been two general categories of approaches that investigators have
used to estimate the degree to which an ecosystem function varies with
microbiome composition: comparative and manipulative. Comparative
studies sample natural variation in an ecosystem function across different
habitats and simultaneously measure variation in community composition.
Investigators can then correlate ecosystem function with aspects of community
composition while attempting to control for environmental variation. These
approaches have
documented important relationships between microbiomes and ecosystem
functions. For example, a meta-analysis of these studies observed a small but
significant contribution of the microbiome to variation in ecosystem
function after controlling for environmental variation [@graham2016]. In
addition, studies focusing on the correlation between the rate of an ecosystem
function and the abundance of an associated functional marker gene observe a
significant correlation, though this relationship is rare and contingent upon
both the function and the ecosystem sampled [@rocca2015]. However, comparative
studies come with some unique challenges and limitations. One issue is that
microbiome attributes tend to covary with the abiotic conditions within an
environment, and whether and how the investigator controls for these variables
can affect the conclusions of such a comparative study. In addition, it is
difficult to know a priori which environmental variables or community attributes
to measure. Finally, while these approaches can establish a potential magnitude
and direction for these relationships, it is often difficult to identify the
taxa or genes that explain the connection between composition and function.

The other broad category of approaches used to address this question are
manipulative approaches. Manipulative experiments try to alter microbial
community composition and observe the effect on function. For example,
reciprocal transplant and common garden experiments have shown that microbiomes
originating from different ecosystems inoculated into the same substrate or
introduced into a common environment display distinct functional rates 
[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009]. In addition,
manipulating diversity by filtering communities by cell size or through dilution
has been shown to alter the rate of ecosystem functions 
[@wagg2014; @wagg2019; @griffiths2001]. However, manipulating the microbiome
directly is challenging. Each of the manipulative approaches applied confound
community composition with other factors. For example, reciprocal transplant and
common garden experiments confound community composition with the abiotic
conditions carried on the inoculum while manipulating composition through
dilution may confound composition with biomass [@reed2007].

(Side note: papers like the Strickland paper were actually asking a subtly
different question from the one that we are asking. Strickland was asking
whether communites can be functionally dissimilar as opposed to functionally
redundant. That is, that historical contingency plays a role in the functioning
of microbial communities. They demonstrated quite well that communities
originating from different environments could have unique functional rates.
However, is that due to microbiome-environment covariance where a different
environment (i.e., a different litter substrate) selects for a litter-degrading
community or is that a result of the unique effect of variation in the
community? They actually point this out in their Discussion, they suggest future
research should determine whether the differences between communities are driven
by taxonomic composition or is function the result of the selection environment
(I think of this is analogous to our "covariation" argument)? They also suggest
trying to adapt communities to a new environment through continuous inputs of
the same resource (just like in our experiment!))

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of whether
variation in the microbiome contributes to variation in the rate of an ecosystem
function. We used artificial ecosystem selection to select for microbiomes that
performed a greater rate of ecosystem function
[@swenson2000; @blouin2015; @panke-buisse2015]. We then tested whether variation
in the microbiome contributed to variation in the rate of ecosystem function and
identified microbiome attributes that might explain this relationship. There are
several potential advantages to this approach for documenting
microbiome-function relationships and for investigating potential mechanisms
underlying those relationships. By passaging microbiomes over multiple rounds
using a common environment, we can weaken the covariance between microbes and
the environment by repeatedly diluting abiotic effects. In addition, by
inoculating our ecosystems with equally sized subsamples, we eliminate the need
to generate microbiome variation through methods that are confounded with
biomass or cell size. Lastly, by comparing our artificially selected community
to a control community resulting from random selection, we can both control for
changes in the environment over time and identify significant genes or taxa
associated with our selection treatment to establish a potential mechanism that
explains this variation.

We applied artificial ecosystem selection to soil microbiomes by selecting on
soil methane oxidation rate. We chose this function because methane is a
globally important greenhouse gas and methane oxidation by soil bacteria is the
primary biological sink for atmospheric methane as well as possibly the main driver
of temporal methane dynamics [@kirschke2013]. In addition, there is evidence that
soil methane oxidation rate may vary with microbiome composition based on
comparative studies in a variety of arctic and tropical ecosystems 
[@meyer2020; @meyer2020a; @freitag2009; @freitag2010] as well as studies that
manipulate methanotroph richness [@schnyder2018]. Finally, methanotrophy is one
of the most deeply conserved microbial physiologies and is represented in a
narrow range of taxa, which suggests that the taxonomic composition of the
microbiome is more likely to be associated with the rate of methane oxidation
than other broader or more shallowly conserved functions
[@schimel1998; @martiny2013].

@@

We estimated the proportion of variation in methane oxidation rate due to
variation in the microbiome as the regression of divergence between the positive
and neutral treatment on the cumulative selection differential [@falconer1996].
Divergence was calculated as the mean methane oxidation rate of the positive
treatment minus the mean methane oxidation rate of the neutral treatment in each
passage. The selection differential was calculated as the difference between
the mean of the three selected jars and the mean of all twelve jars in a passage.
Cumulative selection differential was then the sum of the selection differential
from all preceding selection events. We test this as the regression of
divergence on cumulative selection differential using the 'lm' function.

@@

We estimated the proportion of variation in methane oxidation rate due to
variation in the microbiome by regressing the divergence between the positive
and neutral treatments against the cumulative selection differential. We
observed a slope of 0.40, which would indicate that 40\% of the variation is due
to the microbiome, but this was not significant (statistics).

@@

In this study, we addressed whether variation in the microbiome contributes to
variation in the rate of ecosystem function. To test this, we applied artificial
ecosystem selection to soil microbiomes by selecting on the whole-ecosystem soil
methane oxidation rate. We observed an increase in methane oxidation rate in the
selection treatment relative to the control, which demonstrates that there was a
response to selection. Given that we observed a response to selection, we
conclude that variation in the microbiome contributes to variation in the
methane oxidation rate independent of the environment.

The ultimate goal of this study (and much of microbial ecology) could be
summarized as "Do microbes matter to ecosystem function (or health)?" and there
are at least three ways we could address that question. One would be to ask
whether microorganisms mediate function. This question is perhaps trivial. In
the case of methane emissions from soil, we know that microorganisms produce and
consume methane. We could also ask what would happen if we removed the
microorganisms. Again, this question is trivial. There are no known abiotic
reactions in soil that consume methane (outside of the troposphere, the
stratosphere, and the marine boundary layer @kirschke2013). Therefore, removing
the microbiome would likely eliminate methane production and consumption. Then,
the third question we could ask is whether *variation* in the microbiome is
associated with *variation* in the rate of ecosystem function. The answer to
this question is not so obvious.

If we look at variation in the rate of a microbially-mediated ecosystem function
across a set of habitats, that variation could be associated with variation in
the abiotic environmental conditions, variation in the composition of
microbiomes, or some interaction or covariance between the two (see Figure 5).
To address the question of whether variation in the microbiome matters for
ecosystem function, we need to identify the component of variation uniquely
associated with the microbiome. If we sample natural variation, as in the
comparative approach, it can be difficult to distinguish between the direct
effect of the microbiome and the effect of the microbiome due to covariance with
the environment. Likewise, with the manipulative approach even though we are
placing ecosystem propagules into a common environment or substrate, the
propagule itself carries with it environmental conditions that
cannot be removed. Generating variation in microbiome composition through
filtering or other similar means often confounds composition with other factors,
such as biomass or cell size. To overcome these limitations and to isolate
the direct effect of the microbiome on ecosystem function, we used
artificial ecosystem selection.

To perform artificial ecosystem selection, we generate a set of ecosystems in
bottles using a common, sterile substrate and inoculate them with a propagule
containing a diverse microbiome that is a subsample of an existing microbiome.
This generates variation in microbiome composition among the ecosystems. We can
then measure the rate of an ecosystem function for each bottle and select the
bottles with the highest (or lowest) rate of ecosystem function. The selected
ecosystems are then homogenized and used as the propagule to inoculate a new set of
ecosystems containing the same sterile substrate as the initial set. By
subsampling the microbiome, we can generate variation in the composition of the
microbiome that is independent of cell size or biomass. And by passaging the
microbiome through the same sterile substrate over multiple rounds of selection,
we can progressively minimize the effect of the abiotic conditions from the
donor environment on both the rate of function and the composition of the
microbiome. In this way, over multiple rounds of selection, we can observe
whether varying the microbiome can generate variation in the rate of our
ecosystem function. If we observe no change in ecosystem function from
passage to passage despite our imposed selection, we would conclude that this
ecosystem function in this set of ecosystems is driven primarily by
environmental variation or technical error among the replicates. However, if we
observe a response to selection, that is, a shift in the mean ecosystem function
between successive passages, we would conclude that variation in the microbiome can
generate variation in the rate of ecosystem function. And that is, in fact, what
we observed.

Given that variation in the microbiome is associated with variation in ecosystem
function, a reasonable follow-up question is "How much variation?" This is a
challenging question to answer since the value depends on the population of
ecosystems that is sampled and the estimate for one set of ecosystems may not
apply to another set. For example, we cannot estimate a percent of variation
explained in tropical rainforests and then assume that number applies in polar
wetlands. In addition, the magnitude of the estimate will depend on how much
variation is sampled. A large $\frac{V_M}{V_F}$ is not very interesting if $V_F$
is small.  Keeping these caveats in mind, one way to estimate this number is to
determine how much the recipient jars resemble the selected donor jars that were
used to inoculate them. We can calculate the response to selection as the
difference between two successive passages in their mean methane oxidation rate.
We will denote this as $R$. We can also calculate the strength of selection as
the difference in mean methane oxidation rate between the twelve jars in one
generation and the 3 jars chosen for selection in that generation, which we will
call the selection differential and denote as $S$. If we plot the cumulative
$R$ against the cumulative $S$, the slope of this relationship will equal the
proportion of variation explained by the microbiome.

Using this approach, we observed a relatively large percent of variation
explained by the microbiome, but this number was not significantly different
from zero.

The concept that we have just described is analogous to the concept of
heritability from quantitative genetics. Heritability for given a population is
defined as the proportion of variation in a phenotype associated with variation
in genotype [@lynch1998]. This concept has been applied above the population
level for groups within a species [@goodnight2000], as well as communities and
ecosystems [@whitham2003; @whitham2006]. One study previously estimated the
heritability of a community trait after artificial ecosystem selection using
parent-offspring regression and observed a relatively large heritability of 50\%
[@blouin2015]. However, parent-offspring regression assumes a randomly mating
population, which, by definition, does not apply to a selection experiment. In
the present study, we estimated the variation explained by the microbiome
somewhat differently using the breeder's equation ($R = S h^2$) and regressing
the observed response to selection against the imposed selection differential
(Figure 1; @falconer1996). We also observed a relatively large response to selection
(40\%), though this estimate had large error bars and was not significant so we
did not have the power to estimate an accurate percent of variation explained
for this experiment.

Asking how much of the variation in ecosystem function is due to variation in
the microbiome is a silly question. We cannot answer this questions broadly
for all communities or all ecosystems, because the answer depends in large part
on the magnitude of variation sampled.

What do I want to get across here:

This estimate is analogous to the concept of heritability from quantitative
genetics
Define heritability
Point out that it's been applied to communities and ecosystems
Mention the different ways of estimating heritability
Mention the assumptions inherent in some of these estimates (non-assortative
mating; particularly numbers of shared alleles based on kinship)
Point out that while heritability is a useful framework to borrow from in order
to describe the components of variation, there are limitations in its
application
Propose that we pursue new theory about how to understand what drives variation
in microbiome functions and microbiome-encoded traits

The challenge with applying the concept of heritability outside its original
scope is that it is unclear if the assumptions underlying these estimates apply
to microbiomes. For a genomic trait, we can rely on the assumption that
offspring share, on average, 50\% of the alleles of each of their
parents. It is less clear if offspring share 50\% of the microbes of each of
their parents or if it is sufficient for 50\% of the microbes to be _present_ or
if they must be in a similar relative abundance. Estimates of heritability, such
as parent-offspring regression and the ACE model rely on specific estimates of
the similarity of relatives based on their relatedness. It is unclear if these
assumptions apply to the microbiomes of host ecosystems. It is even less clear
whether they apply to non-host ecosystems that do not reproduce through sexual
recombination.

@@

![Figure 5. Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment. It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome. Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition. For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationshps may also exist.](Figure1.png)
